Evaluation of pilot jobs for Apache Spark\\ applications on HPC clusters

### 2019-08-31 paper.tex

\begin{abstract}
    Big Data has become prominent throughout many scientific fields and, as
    a result, scientific communities have sought out Big Data frameworks to
    accelerate the processing of their increasingly data-intensive
    pipelines. However, while scientific communities typically rely on
    High-Performance Computing (HPC) clusters for the parallelization of
    their pipelines, many popular Big Data frameworks such as Hadoop and
    Apache Spark were primarily designed to be executed on dedicated
    commodity infrastructures. This paper evaluates the benefits of pilot
    jobs over traditional batch submission for Apache Spark on HPC
    clusters. Surprisingly, our results show that the speed-up provided by
    pilot jobs over batch scheduling is moderate to inexistent (0.98 on
    average) despite the presence of long queuing times. In addition, pilot
    jobs provide an extra layer of scheduling that complexifies debugging
    and deployment. We conclude that traditional batch scheduling should
    remain the default strategy to deploy Apache Spark applications on
    HPC clusters.

@@

In contrast to static resource provisioning, pilot jobs are submitted to
the infrastructure separately from the application, to provision resources
on which application tasks will eventually be scheduled. Large pools of
resources can thus be created, shielding applications from the underlying
queuing times, failures, and other idiosyncracies. A variety of frameworks
now rely on pilot jobs, including RADICAL-Pilot~\cite{merzky2015radical},
and recent versions of the Pipeline System for Octave and Matlab
(PSOM)~\cite{bellec2012pipeline}. The survey in~\cite{turilli2018comprehensive} reviews
the current pilot-job systems.

% Goal of the paper
In this paper, we study the use of pilot jobs to run scientific
applications with Apache Spark~\cite{zaharia2016apache} on shared HPC
clusters. The method currently recommended for this purpose involves batch
requesting all the
necessary resources and launching a Standalone Spark cluster once the
resources have been allocated. This is, for instance, the method used on
Compute Canada,
 our national computing infrastructure. With pilot
jobs, rather than requesting all the resources at once, a Spark cluster is
launched with a subset of the resources, and is expanded as more resources
get allocated. Consistently with previous examples of pilot job
deployments, we hypothesize that this strategy would reduce queuing times
by (1) fragmenting resource requirements, and (2) allowing shorter walltime
estimates. While there have been some efforts on implementing pilot jobs
for Apache Spark~\cite{luckow2016hadoop}, research is limited and
none of the them detail their quantitative effect.

Apache Spark is a popular Big Data framework, commonly used in both
industrial and academic settings. Although it is a Scala-based framework,
it also has APIs for Java, Python (PySpark) and R. Spark's Resilient
Distributed Dataset (RDD) abstraction enabled in-memory processing of
pipelines by co-locating tasks and data, which provided important
performance improvements compared to its predecessor Hadoop
MapReduce~\cite{dean2008mapreduce}. Through the use of RDDs, it also became possible
to execute iterative workflows -- something not easily doable in older
frameworks. Schedulers for Spark include its built-in standalone schedule,
Yet Another Resource Negotiator (YARN~\cite{apache13yet}), and Mesos~\cite{hindman2011mesos}. As a
result of the sustained growth in data volumes, Apache Spark and other
Big Data engines are increasingly used for scientific applications,
including in neuroimaging, our primary field of
interest~\cite{boubela2016big,mehta2017comparative,freeman2014mapping}.

@@

    B\'eluga, on the other hand, is a smaller cluster with 872 available
    nodes. Node memory can range between 92 to 752~GB, with the most common
    node type having 186G. All nodes contain 2 x Intel Gold 6148 Skylake @
    2.4~Ghz (40 cores/node) CPU and are connected to each other with a
    100~Gb/s Mellanox Infiniband EDR network. Each non-GPU
    node type contains one 480~GB SSD.

    It is important to note that these clusters were used in production,
    that is, concurrently with other users. This allowed us to test the
    added-value of pilot scheduling in different realistic conditions of queuing
    times. At the time of our experiments, B\'eluga had recently been
    commissioned, resulting in low usage and shorter queuing times overall.
    Conversely, Cedar had been operating for a few years, resulting in
    higher usage and longer queuing times. We repeated our experiments multiple 
    times in each configuration to capture queuing time variability.

@@

    There are three possible cluster managers available to use in Spark applications:
    the Standalone cluster manager, YARN and Mesos. The Standalone cluster manager is the
    most basic cluster manager available and is packaged directly with Spark. YARN is a
    resource manager designed for the resource management of Hadoop applications. However,
    it can be used with different types of applications as well. In contrast, Mesos was 
    designed to manage the resources for a variety of applications. As such, it incorporates
    strategies that are more efficient for use with different frameworks making it possible
    to be used as an HPC scheduler. Due to Standalone's lightweight nature and our intent
    to start Spark clusters per application, we focus on Standalone mode.

@@

    Spark has many configuration options available for dynamic resource provisioning.
    Its standalone cluster provides two deployment modes: client and cluster. The client
    deploy mode executes the driver within the \texttt{spark-submit} process as
    a client to the cluster. The cluster deploy-mode, however, runs the driver within
    a worker process. Such a configuration is practical when the driver program
    is deployed on a machine not co-located in the cluster to reduce network 
    latency between workers and driver. The cluster deploy mode also has the \texttt{supervise}
    feature which allows the driver to be restarted in case of failure.
    This is particularly useful for
    pilot-based overlay clusters where the running driver may fail due to walltime expiration.
    Using the Standalone cluster manager, client mode is supported in all APIs, whereas cluster mode is
    only available for applications using the Scala, Java and R APIs.

@@

    Two different job templates were developed to implement the two main
    conditions compared in our experiments: the batch submission
    template and the pilot submission template. The batch submission template was
    inspired by the template provided by Compute Canada to launch Spark applications 
    on Slurm, \href{https://docs.computecanada.ca/wiki/Apache\_Spark/en}{available here}.
    The template operates as follows: certain resource requirements are requested by
    the user (e.g. walltime, amount of memory per node, number of CPUs per task, number 
    of nodes and number of tasks per node). Once these resources are allocated, a 
    master is started on one of the requested node resources. Then, after the master
    has successfully started, the workers are started on all nodes. Multiple worker
    instances are started on a single node by setting the \texttt{SPARK\_WORKER\_INSTANCES}
    environment variable to the number of tasks per node. Each worker is given as many
     cores as specified by the user in the Slurm resource allocation request.
    After both the masters and the workers have successfully started, the driver is finally
    started. The amount of memory given to each executor corresponds to 95\% of the 
    available memory on the node, to allow for offheap space. 
    The Spark deploy mode selected for the batch template is client mode. We selected this
    deploy mode over cluster mode as driver recovery would not be required in batch in the
    case of walltime expiration, as all the allocated resources will expire at the same time.
    Moreover, many scientific applications are written in Python, which cannot use cluster mode using
    the Standalone resource manager. Should pilot scheduling using cluster mode significantly reduce
    queuing time compared to batch scheduling using client mode, it would provide enough justification to extend
    Spark's Standalone scheduler to include cluster mode for PySpark applications.

    The pilot submission template is similar to that of the batch template, although,
    each pilot will start its own Spark master and worker. However, there will only be one
    pilot which will start the driver process. The pilot selected to start the driver is the first one
    that attempts to do so by way of lockfile.
    The reason for 
    which each pilot starts its own master is to ensure the fault tolerance of the 
    masters. In this configuration, should the active master be killed, one
    of the stand-by masters can takeover and the application may be able to 
    resume if single-node recovery is set. Such a configuration is particularly favourable in pilot scheduling 
    scenarios as node failures may be more frequent due to walltime expiration.
    Additionally, the Spark deploy mode of the driver was selected to be cluster deploy mode.
    This would not only allow the driver to be executed directly on one of the workers,
    but also allow us to make the driver fault-tolerant through the
    \texttt{supervise} mode, which is only available in cluster deploy. As
    with the masters, it is particularly important to have a fault-tolerant
    driver in pilot-scheduling scenarios due to possible walltime
    expiration.
    Should pilots be idle for a certain duration, the
    pilots will terminate themselves such as to not hog resources.
    % \TG{It should be explained before that one of the interests
    % of pilots is to ``play'' with shorter walltimes}. \TG{It would also be
    % nice to have runs where pilots 
    % have shorter walltimes, to see if queuing times are shorter too.}

    Although we start a master on each pilot, we have not incorporated the master recovery function
    for our experiments. This feature is, however, important and should be incorporated in pilot schedulers
    were any nodes can fail by design.

    Due to the differences in deploy modes between batch and pilot submission, 
    batch will always inevitably have one more worker than pilot. This is because
    in cluster deploy, which pilot uses, the driver occupies a worker, whereas in
    client deploy, the driver is separate from any worker.

    Both of these Slurm templates are launched within a Python application called
    \textit{SPA}. The templates are used in conjunction with JSON configuration 
    file and passed to the \href{https://github.com/brentp/slurmpy}{SlurmPy} library within \textit{SPA}.
    The \textit{SPA} application, all the while ensures that all is preconfigured
    correctly before passing it to SlurmPy. It also ensures that enough pilots are
    launched, maintains track of the running/queued pilots, and launches additional
    pilots if there are less pilots than requested by the user in the Slurm queue.

    \subsection{Application}
        \begin{algorithm}\caption{Incrementation}\label{alg:incrementation}              
            \begin{algorithmic}[1]                                                       
                \Input                                                                       
                    \Desc{$x$}{a sleep delay in seconds}                                         
                    \Desc{$n$}{a number of iterations}                                           
                    \Desc{$C$}{a set of image chunks}                                            
                \EndInput                                                                    
                \ForEach{$chunk$ in $C$}                                                      
                    \State read $chunk$                                        
                    \For{$i \in \llbracket 1, n \rrbracket$}                                                         
                        \State $chunk\gets chunk+1$                                              
                        \State sleep $x$                               
                    \EndFor                                                                      
                    \State write $chunk$                                            
                \EndFor                                                                      
            \end{algorithmic}                                                                
        \end{algorithm}
    To determine the added value of pilot scheduling over batch scheduling of Spark
    applications, we required a Spark application operating on a large dataset with
    an important processing time to emulate what would be the average requirements
    of a scientific Spark application. For this, we created a synthetic application 
    that would process the Big Brain~\cite{amunts2013bigbrain}, a 76~GB 3D histological
    image of a human brain. The algorithm is a chain of map transformations that
    at each transformation increment the voxels of the image by 1 (see Algorithm~\ref{alg:incrementation}).
    We chose such a synthetic algorithm
    as the focus of our experiments is pilot scheduling and not the application in
    itself. Furthermore, this algorithm enabled us to have control over the task duration
    which was representative of scientific applications. Additionally, it was important
    that the overall application duration did not vary between the different levels of
    parallelism within our experiments. Being able to adjust the task duration based on
    level of parallelism allowed us to achieve this.

    Spark cluster fault-tolerance is important in determining the suitability of
    pilot-jobs for Spark applications on HPC. Executing jobs on a shared cluster may
    result in a variety of failures. Pilot jobs may be more likely to fail as a result of
    underestimation of resources (e.g. walltimes) and therefore should be fault-tolerant
    to master, worker and driver failures. Although built-in fault-tolerance was not
    investigated in this paper, we want to ensure that the environment is set up in such
    a way that Spark's fault-tolerance configuration could easily be set should our 
    experiments return favourable results for pilot-jobs.
    Fault-tolerance of the driver is only possible in cluster deploy mode, however,
    when using Spark's Standalone scheduler, this mode is not available for Python 
    applications. It is for this reason that our synthetic application is written
    in Scala. Nevertheless, cluster deploy mode is possible for Python application
    using YARN or Mesos schedulers.

@@

    Assuming a divisible load, i.e., the application can be divided in any
    number of tasks, the makespan can be written using the following
    expression, which holds for both batch and pilot execution modes:
    \begin{equation}
        M = \frac{C}{W} \label{eq:mcw}
    \end{equation}
    where:
    \begin{itemize}
        \item $M$ is the makespan of the application
        \item $C$ is the total CPU time of the application
        \item $W$ is the \emph{average} number of Spark workers throughout the execution
    \end{itemize}
    The average number of workers $W$ allows the model to take into account
     variable queuing times. It is computed as follows:
    \begin{equation}
        W = \frac{1}{M}\int_0^M{w(t)dt}\label{eq:avgw}
    \end{equation}
    where $w(t)$ is the number of workers available at time $t$. When the
    application is not subject to any scheduling or queuing time, the
    average number of workers equals the number of workers requested.

    Therefore, assuming a fixed total CPU time, the relation
    between batch and pilot jobs can be represented as:
    \begin{equation}
        \frac{M_{batch}}{M_{pilot}} = \frac{W_{pilot}}{W_{batch}}\label{eq:makespancomp}
    \end{equation}
    where:
    \begin{itemize}
        \item $M_{batch}$ is the makespan of the batch application
        \item $M_{pilot}$ is the makespan of the pilot application
        \item $W_{pilot}$ is the average number of workers of the pilot application
        \item $W_{batch}$ is the average number of workers of the batch application
    \end{itemize}
    We will use this relation to discuss our results later on. It
    corresponds to an ideal case where no data or other overhead is
    present: only queuing times are included, through the integration in
    Equation~\ref{eq:avgw}.

@@

          \hline                                                                           
          \multicolumn{7}{c}{}\\                                                        
        \end{tabular}                                                                    
        \setlength{\belowcaptionskip}{-10pt}                                             
        \caption{Experimental conditions}
        \label{table:conditions}                                                        
        \end{table*}                                                                                   
        To determine if there are any performance benefits to using pilot over 
        batch scheduling, we needed to compare both strategies given various resource
        requirements. It is expected that large batch requests will stay in the 
        resource queue longer than multiple pilot requests that ultimately use up
        the same amount of resources as each individual pilot requests less resources at
        a time. We therefore used four resource configurations to investigate this
        hypothesis (Table~\ref{table:dedicatednodes}).

@@

       As the focus of our experiments here is to measure the impact pilot scheduling
       has on queueing times, we wanted to ensure that our walltime estimates would
       be consistent for each experimental condition. Therefore, we adjusted task 
       delay based on the maximum level of parallelism to ensure that walltimes would
       not need to be readjusted for each configuration. Given utmost parallelism,
       the sleep delay added would amount to a makespan of 1 hour regardless
       of configuration. The actual processing time of the application, however, 
       would differ between configurations as we did not account for task duration (incrementation time, line
       8 in Algorithm~\ref{alg:incrementation})
       in the sleep delay. Due to slight variability in task durations between experiments, in 
       addition to variability in application parallelism due to queuing times of each pilot Slurm job, 
       we have set the walltime to be 2h30 to ensure experiment completion.

@@

    % maybe remove
    %\subsection{Pipeline robustness}
    %    While nodes failures can occur for various reasons, underestimation of 
    %    walltime may lead to early node termination in pilot scheduling systems.
    %    Therefore, it is crucial for the overlay cluster within the pilot system
    %    to be robust to node failures. In this series of experiments, we investigate
    %    Spark's fault-tolerance with respect to worker, master and driver termination.

@@

As can be seen in both Figures~\ref{fig:makespansbeluga} and~\ref{fig:makespanscedar}, pilots 
generally did not bring any performance improvement to the execution of Spark workloads
on either system. In B\'eluga and Cedar, 16 pilots were generally slower than batch except for a 
few occurrencs. When 16 pilots were faster, it was generally either very slight, or, as expected,
when batch queuing times were very large. The maximum speedup achieved
was approximately 5.6~x for 8 and 16 pilots on B\'eluga and 2.6~x on Cedar. 
It is only in Configuration~3 (Figure~\ref{fig:beluga4}), on B\'eluga,
that we spot any large speedups, however, it generally remains that
pilots are slightly slower, even in Configuration~3. The same can be said about Cedar, however,
it occured with Configuration~4 instead of 3. On average, pilots were up to 0.11~x slower on B\'eluga and
0.17~x slower on Cedar (see Table~\ref{table:speedup}), both of which occurred when using 16 pilots.
The only times when pilots were found to be, on average, faster, was in Configuration~3 on B\'eluga
and Configuration~4 on Cedar where the speedup was found to reach a max of 1.46 for 8 pilots on 
B\'eluga. The missing bars in Figures~\ref{fig:makespansbeluga} and~\ref{fig:makespanscedar} correspond to 
failed executions.

@@

In the case of pilots, multiple Slurm jobs, each with potentially different queuing times,
were used. Figure~\ref{fig:mwall} displays the makespan times for an average number of workers
given an experiment configuration, where the coloured line corresponds to the estimated makespan using 
our performance model. Makespan for the model was calculated as:
$$
M = \frac{125\times(d + 20)}{W}\times 10,
$$
where:
\begin{itemize}
    \item $125$ is the number of chunks BigBrain was split into
    \item $M$ is the makespan of the application, in seconds
    \item $W$ is the average number of Spark workers throughout the execution, computed as in Equation~\ref{eq:avgw}
    \item $d$ is the task delay, in seconds, associated with a given configuration
    \item $20$ is the average measured incrementation duration, in seconds, for a BigBrain chunk
    \item $10$ is the number of iterations
\end{itemize}
As can be seen in the Figure~\ref{fig:mwall}, both pilots and batch 
followed the general trend denoted by the model, which confirms that 
the system was behaving without major external sources of
overhead. However, pilots were consistently slower than batch for the same
number of average workers. This is particularly visible in Configuration 4
on B\'eluga (Fig. \ref{fig:mwbeluga}), and in Configuration 2, 3 and 4 on
Cedar (Fig. \ref{fig:mwcedar}). It means that some sort of overhead
impacted pilots but not batch. It can also be noticed that both batch and
pilots occasionally deviated from the model line, in particular
Configurations 2 and 3 in B\'eluga: this is due to the fact that our
application is not a divisible load, as the model assumes. Extending the
model beyond divisible loads is easy enough for batch and confirms that the
observed deviations from the model line come from this assumption (see gray
line in Fig.~\ref{fig:mwall}). The extended expression for pilots is more complex
though.

@@

Pilots are meant to allow for underestimation of resources, particularly
walltime, which would not be known to the user prior to execution. In its
current state, \texttt{SPA} is not able to evaluate that functionality.
Further experiments on master fault-tolerance, driver recovery, and
checkpointing will need to be conducted to determine if pilot jobs are even
a viable solution for Spark applications running on HPC.

>>>>

\begin{abstract}
    Big Data has become prominent throughout many scientific fields and, as
    a result, scientific communities have sought out Big Data frameworks to
    accelerate the processing of their increasingly data-intensive
    pipelines. However, while scientific communities typically rely on
    High-Performance Computing (HPC) clusters for the parallelization of
    their pipelines, many popular Big Data frameworks such as Hadoop and
    Apache Spark were primarily designed to be executed on dedicated
    commodity infrastructures. This paper evaluates the benefits of pilot
    jobs over traditional batch submission for Apache Spark on HPC
    clusters. Surprisingly, our results show that the speed-up provided by
    pilot jobs over batch scheduling is moderate to non-existent (0.98 on
    average) despite the presence of long queuing times. In addition, pilot
    jobs provide an extra layer of scheduling that complicates debugging
    and deployment. We conclude that traditional batch scheduling should
    remain the default strategy to deploy Apache Spark applications on
    HPC clusters.

@@

In contrast to static resource provisioning, pilot jobs are submitted to
the infrastructure separately from the application, to provision resources
on which application tasks will eventually be scheduled. Large pools of
resources can thus be created, shielding applications from the underlying
queuing times, failures, and other idiosyncrasies. A variety of frameworks
now rely on pilot jobs, including RADICAL-Pilot~\cite{merzky2015radical},
and recent versions of the Pipeline System for Octave and Matlab
(PSOM)~\cite{bellec2012pipeline}. The survey in~\cite{turilli2018comprehensive} reviews
the current pilot-job systems.

% Goal of the paper
In this paper, we study the use of pilot jobs to run scientific
applications with Apache Spark~\cite{zaharia2016apache} on shared HPC
clusters. The currently recommended method for this purpose involves batch
requesting all the
necessary resources and launching a Standalone Spark cluster once the
resources have been allocated. This is, for instance, the method used on
Compute Canada,
 our national computing infrastructure. With pilot
jobs, rather than requesting all the resources at once, a Spark cluster is
launched with a subset of the resources, and is expanded as more resources
get allocated. Consistently with previous examples of pilot job
deployments, we hypothesize that this strategy would reduce queuing times
by (1) fragmenting resource requirements, and (2) allowing shorter wall-time
estimates. While there have been some efforts on implementing pilot jobs
for Apache Spark~\cite{luckow2016hadoop}, research is limited and
none of the them detail their quantitative effect.

Apache Spark is a popular Big Data framework, commonly used in both
industrial and academic settings. Although it is a Scala-based framework,
it also has APIs for Java, Python (PySpark) and R. Spark's Resilient
Distributed Dataset (RDD) abstraction enabled in-memory processing of
pipelines by co-locating tasks and data, which provided important
performance improvements compared to its predecessor Hadoop
MapReduce~\cite{dean2008mapreduce}. Through the use of RDDs, it also became possible
to execute iterative workflows -- something not easily doable in older
frameworks. Schedulers for Spark include its built-in Standalone scheduler,
Yet Another Resource Negotiator (YARN~\cite{apache13yet}), and Mesos~\cite{hindman2011mesos}. As a
result of the sustained growth in data volumes, Apache Spark and other
Big Data engines are increasingly used for scientific applications,
including in neuroimaging, our primary field of
interest~\cite{boubela2016big,mehta2017comparative,freeman2014mapping}.

@@

    B\'eluga, on the other hand, is a smaller cluster with 872 available
    nodes. Node memory can range between 92 to 752~GB, with the most common
    node type having 186~GB. All nodes contain 2 x Intel Gold 6148 Skylake @
    2.4~Ghz (40 cores/node) CPU and are connected to each other with a
    100~Gb/s Mellanox Infiniband EDR network. Each non-GPU
    node type contains one 480~GB SSD.

    It is important to note that these clusters were used in production.
    That is, concurrently with other users. This allowed us to test the
    added-value of pilot scheduling in different realistic conditions of queuing
    times. At the time of our experiments, B\'eluga had recently been
    commissioned, resulting in low usage and shorter queuing times overall.
    Conversely, Cedar had been operating for a few years, resulting in
    higher usage and longer queuing times. We repeated our experiments multiple 
    times in each configuration to capture queuing time variability.

@@

    There are three possible cluster managers available to use with Spark applications:
    the Standalone cluster manager, YARN and Mesos. The Standalone cluster manager is the
    most basic cluster manager available and is packaged directly with Spark. The 
    YARN resource manager is designed for the resource management of Hadoop applications, however,
    it can be used with different types of applications as well. In contrast, Mesos was 
    designed to manage the resources for a variety of applications. As such, it incorporates
    strategies that are more efficient for use with different frameworks making it possible
    to be used as an HPC scheduler. Due to Standalone's lightweight nature and our intent
    to start Spark clusters per application, we utilized the Standalone cluster manager.

@@

    Spark has many configuration options available for dynamic resource provisioning.
    Its standalone cluster provides two deployment modes: client and cluster. The client
    deploy mode executes the driver within the \texttt{spark-submit} process as
    a client to the cluster. The cluster deploy-mode, however, runs the driver within
    a worker process. Such a configuration is practical when the driver program
    is deployed on a machine not co-located in the cluster to reduce network 
    latency between workers and the driver. The cluster deploy mode also has the \texttt{supervise}
    feature which allows the driver to be restarted in case of failure.
    This is particularly useful for
    pilot-based overlay clusters where the running driver may fail due to wall-time expiration.
    Using the Standalone cluster manager, client mode is supported in all APIs, whereas cluster mode is
    only available for applications using the Scala, Java and R APIs.

@@

    Two different job templates were developed to implement the two main
    conditions compared in our experiments: the batch submission
    template and the pilot submission template. The batch submission template was
    inspired by the template provided by Compute Canada to launch Spark applications 
    on Slurm, \href{https://docs.computecanada.ca/wiki/Apache\_Spark/en}{available here}.
    The template operates as follows: certain resource requirements are requested by
    the user (e.g. wall-time, amount of memory per node, number of CPUs per task, number 
    of nodes and number of tasks per node). Once these resources are allocated, a 
    master is started on one of the requested node resources. Then, after the master
    has successfully started, the workers are started on all nodes. Multiple worker
    instances are started on a single node by setting the \texttt{SPARK\_WORKER\_INSTANCES}
    environment variable to the number of tasks per node. Each worker is given as many
     cores as specified by the user in the Slurm resource allocation request.
    After both the masters and the workers have successfully started, the driver is finally
    started. The amount of memory given to each executor corresponds to 95\% of the 
    available memory on the node, to allow for offheap space. 
    The Spark deploy mode selected for the batch template is client mode. We selected this
    deploy mode over cluster mode as driver recovery would not be required in batch in the
    case of wall-time expiration, as all the allocated resources will expire at the same time.
    Moreover, many scientific applications are written in Python, which cannot use cluster mode using
    the Standalone resource manager. Should pilot scheduling using cluster mode significantly reduce
    queuing time compared to batch scheduling using client mode, it would provide enough justification to extend
    Spark's Standalone scheduler to include cluster mode for PySpark applications.

    The pilot submission template is similar to that of the batch template, although,
    each pilot will start its own Spark master and worker. However, there will only be one
    pilot which will start the driver process. The pilot selected to start the driver is the first one
    that attempts to do so by means of a lockfile.
    The reason for 
    which each pilot starts its own master is to ensure the fault tolerance of the 
    masters. In this configuration, should the active master be killed, one
    of the stand-by masters can takeover and the application may be able to 
    resume if single-node recovery is set. Such a configuration is particularly favourable in pilot scheduling 
    scenarios as node failures may be more frequent due to wall-time expiration.
    Additionally, the Spark deploy mode of the driver was selected to be cluster deploy mode.
    This would not only allow the driver to be executed directly on one of the workers,
    but also allow us to make the driver fault-tolerant through the
    \texttt{supervise} mode, which is only available in cluster deploy. As
    with the masters, it is particularly important to have a fault-tolerant
    driver in pilot-scheduling scenarios due to possible wall-time
    expiration.
    Should pilots be idle for a certain duration, the
    pilots will terminate themselves in order to not hog resources.
    % \TG{It should be explained before that one of the interests
    % of pilots is to ``play'' with shorter wall-times}. \TG{It would also be
    % nice to have runs where pilots 
    % have shorter wall-times, to see if queuing times are shorter too.}

    Although we start a master on each pilot, we have not incorporated the master recovery function
    for our experiments. However, this feature is important and should be incorporated in pilot schedulers
    where any nodes can fail by design.

    Due to the differences in deploy modes between batch and pilot submission, 
    batch will always inevitably have one more worker than pilot. This is because
    in cluster deploy, which pilot uses, the driver occupies a worker, whereas in
    client deploy, the driver is a distinct process.

    Both of these Slurm templates are launched within a Python application called
    \textit{SPA}. The templates are used in conjunction with JSON configuration 
    file and passed to the \href{https://github.com/brentp/slurmpy}{SlurmPy} library within \textit{SPA}.
    The \textit{SPA} application ensures that all is preconfigured
    correctly before passing it to SlurmPy. It also ensures that enough pilots are
    launched, maintains track of the running/queued pilots, and launches additional
    pilots if there are less pilots than requested by the user in the Slurm queue.

    \subsection{Application}
        \begin{algorithm}\caption{Incrementation}\label{alg:incrementation}              
            \begin{algorithmic}[1]                                                       
                \Input                                                                       
                    \Desc{$x$}{a sleep delay in seconds}                                         
                    \Desc{$n$}{a number of iterations}                                           
                    \Desc{$C$}{a set of image chunks}                                            
                \EndInput                                                                    
                \ForEach{$chunk$ in $C$}                                                      
                    \State read $chunk$                                        
                    \For{$i \in \llbracket 1, n \rrbracket$}                                                         
                        \State $chunk\gets chunk+1$                                              
                        \State sleep $x$                               
                    \EndFor                                                                      
                    \State write $chunk$                                            
                \EndFor                                                                      
            \end{algorithmic}                                                                
        \end{algorithm}
    To determine the added value of pilot over batch scheduling of Spark
    applications, we required a Spark application operating on a large dataset with
    an important processing time to emulate what would be the average requirements
    of a scientific Spark application. For this, we created a synthetic application 
    that would process the 40~$\mu$m BigBrain~\cite{amunts2013bigbrain}, a 76~GB 3D histological
    image of a human brain. The algorithm is a chain of map transformations that,
    at each transformation, increment the voxels of the image by 1 (see Algorithm~\ref{alg:incrementation}).
    We selected a synthetic algorithm
    as the focus of our experiments is pilot scheduling and not the application in
    and of itself. Furthermore, this algorithm enabled us to have control over the task duration and make it representative of scientific applications. Additionally, it was important
    that the overall application duration did not vary between the different levels of
    parallelism within our experiments. Being able to adjust the task duration based on
    level of parallelism allowed us to achieve this.

    Spark cluster fault-tolerance is important in determining the suitability of
    pilot-jobs for Spark applications on HPC. Executing jobs on a shared cluster may
    result in a variety of failures. Pilot jobs may be more likely to fail as a result of
    underestimation of resources (e.g. wall-times), and therefore, should be fault-tolerant
    to master, worker and driver failures. Although built-in fault-tolerance was not
    investigated in this paper, we want to ensure that the environment is set up in such
    a way that Spark's fault-tolerance configuration could easily be set should our 
    experiments return favourable results for pilot-jobs.
    Fault-tolerance of the driver is only possible in cluster deploy mode, however,
    when using Spark's Standalone scheduler, this mode is not available for Python 
    applications. It is for this reason that our synthetic application is written
    in Scala. Nevertheless, cluster deploy mode is available for PySpark applications
    using YARN or Mesos schedulers.

@@

    Assuming a divisible load, i.e., the application can be divided in any
    number of tasks, the makespan can be written using the following
    expression, which holds for both batch and pilot execution modes:
    \begin{equation}
        M = \frac{C}{W} \label{eq:mcw}
    \end{equation}
    where:
    \begin{itemize}
        \item $M$ is the makespan of the application
        \item $C$ is the total wall-time of the application
        \item $W$ is the \emph{average} number of Spark workers throughout the execution
    \end{itemize}
    The average number of workers $W$ allows the model to take into account
     variable queuing times. It is computed as follows:
    \begin{equation}
        W = \frac{1}{M}\int_0^M{w(t)dt}\label{eq:avgw}
    \end{equation}
    where $w(t)$ is the number of workers available at time $t$. When the
    application is not subject to any scheduling or queuing time, the
    average number of workers equals the number of workers requested.

    Therefore, assuming a fixed total wall-time, the relation
    between batch and pilot jobs can be represented as:
    \begin{equation}
        \frac{M_{batch}}{M_{pilot}} = \frac{W_{pilot}}{W_{batch}}\label{eq:makespancomp}
    \end{equation}
    where:
    \begin{itemize}
        \item $M_{batch}$ is the makespan of the batch application
        \item $M_{pilot}$ is the makespan of the pilot application
        \item $W_{pilot}$ is the average number of workers of the pilot application
        \item $W_{batch}$ is the average number of workers of the batch application
    \end{itemize}
    We will use this relation to discuss our results later on. It
    corresponds to an ideal case where no data or other overhead is
    present: only queuing times are included, through the integration in
    Equation~\ref{eq:avgw}.

@@

          \hline                                                                           
          \multicolumn{7}{c}{}\\                                                        
        \end{tabular}                                                                    
        \setlength{\belowcaptionskip}{-10pt}                                             
        \caption{Experimental conditions}
        \label{table:conditions}                                                        
        \end{table*}                                                                                   
        To determine if there are any performance benefits to using pilot over 
        batch scheduling, we needed to compare both strategies given various resource
        requirements. It is expected that large batch requests will stay in the 
        resource queue longer than multiple pilot requests that ultimately use up
        the same amount of resources, as each individual pilot requests fewer resources at
        a time. We therefore used four resource configurations to investigate this
        hypothesis (Table~\ref{table:dedicatednodes}).

@@

       As the focus of our experiments is to measure the impact pilot scheduling
       has on queueing times, we wanted to ensure that our wall-time estimates would
       be consistent for each experimental condition. Therefore, we adjusted task 
       delay based on the maximum level of parallelism to ensure that wall-times would
       not need to be readjusted for each configuration. Given utmost parallelism,
       the sleep delay added would amount to a makespan of 1 hour regardless
       of configuration. The actual processing time of the application, however, 
       would differ between configurations as we did not account for task duration (incrementation time, line
       8 in Algorithm~\ref{alg:incrementation})
       in the sleep delay. Due to slight variability in task durations between experiments, in 
       addition to variability in application parallelism due to queuing times of each pilot Slurm job, 
       we set the Slurm allocation wall-time to be 2h30 to ensure experiment completion.

@@

    % maybe remove
    %\subsection{Pipeline robustness}
    %    While nodes failures can occur for various reasons, underestimation of 
    %    wall-time may lead to early node termination in pilot scheduling systems.
    %    Therefore, it is crucial for the overlay cluster within the pilot system
    %    to be robust to node failures. In this series of experiments, we investigate
    %    Spark's fault-tolerance with respect to worker, master and driver termination.

@@

As can be seen in both Figures~\ref{fig:makespansbeluga} and~\ref{fig:makespanscedar}, pilots 
generally did not bring any performance improvement to the execution of Spark workloads
on either system. In B\'eluga and Cedar, 16 pilots were generally slower than batch except for a 
few occurrences. When 16 pilots were faster, it was generally either very slight, or, as expected,
when batch queuing times were very large. The maximum speedup achieved
was approximately 5.6~x for 8 and 16 pilots on B\'eluga and 2.6~x on Cedar. 
It is only in Configuration~3 (Figure~\ref{fig:beluga4}) on B\'eluga
that we spot any large speedups, however, it generally remains that
pilots are slightly slower, even in this configuration. The same can be said about Cedar, however,
it occurred with Configuration~4 instead of 3. On average, pilots were up to 0.11~x slower on B\'eluga and
0.17~x slower on Cedar (see Table~\ref{table:speedup}), both of which occurred when using 16 pilots.
The only times when pilots were found to be, on average, faster, was in Configuration~3 on B\'eluga
and Configuration~4 on Cedar where the speedup was found to reach a max of 1.46 for 8 pilots on 
B\'eluga. The missing bars in Figures~\ref{fig:makespansbeluga} and~\ref{fig:makespanscedar} correspond to 
failed executions.

@@

In the case of pilots, multiple Slurm jobs, each with potentially different queuing times,
were used. Figure~\ref{fig:mwall} displays the makespan times for an average number of workers
given an experiment configuration, where the coloured line corresponds to the estimated makespan using 
our performance model. Makespan for the model was calculated as:
$$
M = \frac{125\times(d + 20)}{W}\times 10,
$$
where:
\begin{itemize}
    \item $125$ is the number of chunks BigBrain was split into
    \item $M$ is the makespan of the application, in seconds
    \item $W$ is the average number of Spark workers throughout the execution, computed as in Equation~\ref{eq:avgw}
    \item $d$ is the task delay, in seconds, associated with a given configuration
    \item $20$ is the average measured incrementation duration, in seconds, for a BigBrain chunk
    \item $10$ is the number of iterations
\end{itemize}
As can be seen in the Figure~\ref{fig:mwall}, both pilots and batch 
followed the general trend denoted by the model, which confirms that 
the system was behaving without major external sources of
overhead. However, pilots were consistently slower than batch for the same
number of average workers. This is particularly visible in Configuration 4
on B\'eluga (Fig. \ref{fig:mwbeluga}), and in Configuration 2, 3 and 4 on
Cedar (Fig. \ref{fig:mwcedar}). It means that some sort of overhead
impacted pilots but not batch. It can also be noticed that both batch and
pilots occasionally deviated from the model line, in particular
Configurations 2 and 3 in B\'eluga: this is due to the fact that our
application is not a divisible load, as the model assumes. Extending the
model beyond divisible loads is easy enough for batch and confirms that the
observed deviations from the model line come from this assumption (see grey
line in Fig.~\ref{fig:mwall}). The extended expression for pilots is more complex
though.

@@

Pilots are meant to allow for underestimation of resources, particularly
wall-time, which would not be known to the user prior to execution. In its
current state, \texttt{SPA} is not able to evaluate that functionality.
Further experiments on master fault-tolerance, driver recovery, and
checkpointing will need to be conducted to determine if pilot jobs are even
a viable solution for Spark applications running on HPC.
